---
title: "Network and VIP Analysis"
author: "Your Name"
date: "2024-06-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Required Libraries
Load the necessary libraries for data processing and analysis.

```{r load_libraries}
library(dplyr)
library(multinet)
library(ggplot2)
```

## Functions for Network Analysis
Define functions to process network data and obtain layer counts.

```{r functions_network_analysis}
get_num_layers <- function(net, node_id, neighbors, start_year, end_year) {
  results <- numeric(length(neighbors))
  names(results) <- neighbors
  
  for (year in start_year:end_year) {
    neigh <- neighbors_ml(net, node_id, layers = year)
    for (n in neigh) {
      if (n %in% neighbors) {
        results[n] <- results[n] + 1
      }
    }
  }
  
  output <- data.frame(node_id = node_id, num_layers = results, neighbor_id = names(results))
  output$neighbor_id <- sub("^[^.]+//.", "", output$neighbor_id)
  return(output)
}

process_node_ids <- function(net, node_ids, start_year, end_year) {
  results_list <- list()
  
  for (node_id in node_ids) {
    # Check if node_id is in the network
    if (!node_id %in% actors_ml(net)$actor) {
      message(paste("Node", node_id, "is not in the network"))
      next
    }
    
    neighbors <- neighbors_ml(net, actor = node_id)
    results_list[[node_id]] <- get_num_layers(net, node_id, neighbors, start_year, end_year)
  }
  
  result_df <- bind_rows(results_list)
  colnames(result_df) <- c("node_id", "num_layers", "neighbor_id")
  return(result_df)
}
```

## Processing Initial Nodes and Their Neighbors
Process the initial set of nodes and then the neighbors of those nodes.

```{r process_nodes}
# Read in the multinet network file
network_file <- "../visualization/Input/multinet_network_input_bog.txt"
net <- read_ml(network_file, aligned = TRUE)

# Process initial node_ids
initial_node_ids <- c("PLGY01")
test <- process_node_ids(net, initial_node_ids, 2011, 2017)

# Process neighbors of initial node
neighbors <- neighbors_ml(net, "PLGY01")
test_neighbors <- process_node_ids(net, neighbors, 2011, 2017)

# Concatenate test and test_neighbors
combined_df <- bind_rows(test, test_neighbors)

# Create a consistent identifier for each pair to identify duplicates
unique_combined_df <- combined_df %>%
  mutate(pair_id = pmin(node_id, neighbor_id),
         pair_id_2 = pmax(node_id, neighbor_id)) %>%
  distinct(pair_id, pair_id_2, .keep_all = TRUE) %>%
  select(-pair_id, -pair_id_2)

# Print the unique_combined_df to check the results
print(unique_combined_df)

m_stor_neighbors_unique <- unique_combined_df
```

## Loading VIP Values
Load VIP values for different parameters from text files.

```{r load_vip_values}
vip_values_N_path <- "Input/Bog.08.VIP_values_with_N.percent__blue.txt"
vip_values_CO2_path <- "Input/Bog.08.VIP_values_with_CO2.mM__blue.txt"
vip_values_alphaC_path <- "Input/Bog.08.VIP_values_with_alphaCblue.txt"
vip_values_d13C_path <- "Input/Bog.08.VIP_values_with_d13C_CO2__blue.txt"

# Function to load VIP values
load_vip_values <- function(file_path, parameter) {
  data <- read.table(file_path, header = FALSE, sep = " ", fill = TRUE)
  data <- data %>% select(V5, V10) %>% rename(ID = V5, VIP = V10)
  data$Parameter <- parameter
  return(data)
}

# Load data
df_N <- load_vip_values(vip_values_N_path, "N percent")
df_CO2 <- load_vip_values(vip_values_CO2_path, "CO2 mM")
df_alphaC <- load_vip_values(vip_values_alphaC_path, "alphaC")
df_d13C <- load_vip_values(vip_values_d13C_path, "d13C CO2")

# Combine data
combined_df <- bind_rows(df_N, df_CO2, df_alphaC, df_d13C)
```

## Joining Network Data with VIP Values
Join the network analysis results with the VIP values for both nodes and their neighbors.

```{r join_vip_values}
combined_data <- m_stor_neighbors_unique %>%
  left_join(combined_df, by = c("node_id" = "ID")) %>%
  rename(VIP_node = VIP, Parameter_node = Parameter) %>%
  left_join(combined_df, by = c("neighbor_id" = "ID")) %>%
  rename(VIP_neighbor = VIP, Parameter_neighbor = Parameter) %>%
  distinct()

# Display the first few rows of the result
head(combined_data)
```

## Filtering the Combined Data
Filter the combined data just for M.stor's neighbors.

```{r filter_data}
filtered_data <- combined_data %>%
  filter((node_id == "PLGY01" | neighbor_id == "PLGY01") & VIP_neighbor > 1 & num_layers >= 6)

# Display the first few rows of the filtered result
unique(filtered_data$neighbor_id)
```

```{r}

filtered_data_neighbors <- filtered_data %>%
  select(-node_id, -VIP_node, -Parameter_node, -Parameter_neighbor, -VIP_neighbor) %>% 
  unique() %>%
  add_row(neighbor_id = "PLGY01", num_layers = NA)


```

```{r}

# Load necessary libraries
library(readxl)

# Load the user's uploaded Excel file
file_path <- "Input/all_vip_env_var_corr_DSedits.xlsx"

# Read the first sheet of the Excel file
vip_env_var_corr_df <- read_excel(file_path, sheet = 1)

# Filter data to include only rows in the blue module
blue_module_df <- subset(vip_env_var_corr_df, `Module Color` == 'blue')

# Perform a merge on 'neighbor_id' from filtered_data_neighbors and 'genome' from blue_module_df,
# and also ensure 'Parameter_neighbor' matches 'Environmental Variable'
# Perform a merge on 'neighbor_id' from filtered_data_neighbors and 'genome' from blue_module_df
merged_df <- merge(filtered_data_neighbors, blue_module_df, 
                   by.x = 'neighbor_id', 
                   by.y = 'genome', 
                   all.x = TRUE) %>% unique()

# Display the first few rows of the merged dataframe
print(head(merged_df))


```

```{r}
# Define y_limits and custom_colors for this chunk's plot
y_limits <- c(0, 5)
custom_colors <- c(
  "d13C CO2" = "#E06666", 
  "alphaC" = "#93C47D", 
  "[CO2] (mM)" = "#F6B26B", 
  "CO2 mM" = "#F6B26B",
  "N percent" = "#76A5AF",
  "N.percent" = "#76A5AF",
  "d13C_CO2" = "#E06666",
  "CO2.mM" = "#F6B26B"
)

# Function to create the plot for each neighbor with or without labels
create_vip_correlation_plot <- function(df, id, show_labels = FALSE) {
  df_id <- df %>% filter(ID == id)
  
  # Set consistent factor levels
  df_id$Parameter <- factor(df_id$Parameter, levels = c("CO2.mM", "N.percent", "alphaC"))
  
  # Create the plot
  p <- ggplot(df_id) +
    geom_bar(aes(x = as.numeric(Parameter) - 0.2, y = VIP, fill = Parameter), 
             stat = "identity", width = 0.4, alpha = 0.6) +
    geom_point(aes(x = as.numeric(Parameter) + 0.2, y = Correlation_scaled, color = Parameter), 
               size = 12) +  # Large size for correlation points
    geom_text(aes(x = as.numeric(Parameter), y = Correlation_scaled + 0.25, 
                  label = ifelse(PValue < 0.05, "*", ""), color = Parameter),
              vjust = 0, size = 8) +  # Adjust size and vertical position for the significance star
    geom_hline(aes(yintercept = 1, linetype = "VIP Threshold"), color = "blue", linewidth = 2.0) +
    geom_hline(aes(yintercept = 0.5 * 5, linetype = "Correlation Threshold"), color = "red", linewidth = 2.0) +
    scale_x_continuous(
      breaks = 1:3, 
      labels = if (show_labels) c("CO2 mM", "%N", "alphaC") else NULL,
      expand = c(0.1, 0.1)
    ) +
    scale_y_continuous(name = NULL, limits = y_limits,  # Extend y-axis limit
                       sec.axis = sec_axis(~ . / 5, name = NULL, breaks = seq(0, 1, by = 0.2))) +
    scale_linetype_manual(name = "Thresholds", values = c("VIP Threshold" = "dashed", "Correlation Threshold" = "dotted")) +
    labs(x = NULL, title = NULL) +  # Remove figure title and x-axis label
    scale_fill_manual(values = custom_colors) +
    scale_color_manual(values = custom_colors)
  
  # Conditional theme: if show_labels is TRUE, show axis numbers and legend, otherwise hide them
  if (show_labels) {
    p <- p + theme_minimal() +
      theme(
        legend.position = "bottom",  # Legend at the bottom
        legend.box = "vertical",     # Stack legend vertically if needed
        legend.margin = margin(t = 20),  # Add margin above the legend
        axis.line = element_line(linewidth = 1.5, color = "black"),
        axis.title.y = element_blank(),  # No axis titles
        axis.title.y.right = element_blank(),  # No right y-axis title
        axis.text.x = element_text(size = 15, color = "black"),
        axis.text.y = element_text(size = 15, color = "black"),
        axis.ticks.x = element_line(linewidth = 1.5, color = "black"),
        axis.ticks.length.x = unit(0.3, "cm"),
        axis.ticks.y = element_line(linewidth = 1.5, color = "black"),
        plot.margin = margin(t = 20, r = 20, b = 50, l = 20)  # Increase bottom margin to fit legend
      )
  } else {
    p <- p + theme_minimal() +
      theme(
        legend.position = "none",
        axis.line = element_line(linewidth = 1.5, color = "black"),  # Keep axis lines
        axis.title = element_blank(),  # No axis titles
        axis.text = element_blank(),  # Remove axis text (numbers)
        axis.ticks.x = element_line(linewidth = 1.5, color = "black"),
        axis.ticks.length.x = unit(0.3, "cm"),
        axis.ticks.y = element_line(linewidth = 1.5, color = "black"),
        plot.margin = margin(t = 10, r = 10, b = 10, l = 10)  # Smaller margins for compact plots
      )
  }
  
  return(p)
}
# Prepare plot_data by renaming and calculating necessary columns
plot_data <- merged_df %>%
  rename(
    ID = neighbor_id,
    Parameter = `Environmental Variable`,
    VIP = `VIP Score`,
    PValue = p.value
  ) %>%
  mutate(Correlation_scaled = spearman_correlation * 5)  # Scale correlation as in the plotting function

# Check the structure of plot_data
str(plot_data)

# Create and print plots for all unique neighbor IDs in the dataset
for (id in unique(plot_data$ID)) {
  if (id == "PLGY01") {
    plot <- create_vip_correlation_plot(plot_data, id, show_labels = TRUE)  # Show labels for PLGY01
  } else {
    plot <- create_vip_correlation_plot(plot_data, id, show_labels = FALSE)  # Hide labels for other plots
  }
  print(plot)
}



```

```{r}
# Prepare plot_data by renaming columns and using raw correlation values
plot_data <- merged_df %>%
  rename(
    ID = neighbor_id,
    Parameter = `Environmental Variable`,
    VIP = `VIP Score`,
    Correlation = spearman_correlation  # Use raw correlation values
  )

# Check the structure of plot_data
str(plot_data)

# Calculate the maximum y-limit across the entire dataset for VIP values
y_limits <- c(0, max(plot_data$VIP) * 1.2)

create_vip_correlation_plot <- function(df, id, show_labels = FALSE) {
  df_id <- df %>% filter(ID == id)
  
  # Define the colors for VIP and Correlation
  vip_color <- "#E07B91"  # Color for VIP values
  correlation_color <- "#40B0A6"  # Color for Correlation values
  
  # Order the Parameter factor based on consistent levels
  df_id$Parameter <- factor(df_id$Parameter, levels = c("CO2.mM", "N.percent", "alphaC"))
  
  # Transform correlation values to match the VIP axis limits
  df_id <- df_id %>%
    mutate(Correlation_scaled = Correlation * max(y_limits))  # Scale correlation to match y_limits

  # Create the plot
  p <- ggplot(df_id) +
    # Square for VIP values
    geom_point(aes(x = as.numeric(Parameter) - 0.2, y = VIP), 
               shape = 15, color = vip_color, size = 15, alpha = 0.6) +
    # Circle for scaled correlation values
    geom_point(aes(x = as.numeric(Parameter) + 0.2, y = Correlation_scaled), 
               shape = 16, color = correlation_color, size = 15) +
    # Dashed horizontal lines
    geom_hline(yintercept = 1, color = vip_color, linetype = "dashed", linewidth = 2.0) +
    geom_hline(yintercept = 0.5 * max(y_limits), color = correlation_color, linetype = "dashed", linewidth = 2.0) +
    scale_x_continuous(
      breaks = 1:3, 
      labels = if (show_labels) c("CO2 mM", "%N", "alphaC") else NULL,  # Show x-axis labels conditionally
      expand = c(0.1, 0.1)
    ) +
    scale_y_continuous(
      name = if (show_labels) "VIP Value" else NULL, limits = y_limits,  
      sec.axis = sec_axis(~ . / max(y_limits), name = if (show_labels) "Spearman's rho" else NULL, breaks = seq(0, 1, by = 0.2))  # Fixed 0-1 range for correlation
    ) +
    labs(x = NULL) +  # Removed the title
    theme_minimal() +
    theme(
      panel.grid = element_blank(),  # Remove background grid
      legend.position = "none",
      axis.line = element_line(size = 1.5, color = "black"),
      axis.line.y.left = element_line(color = vip_color, size = 1.5),  # Color VIP y-axis line
      axis.line.y.right = element_line(color = correlation_color, size = 1.5),  # Color Correlation y-axis line
      axis.text.x = if (show_labels) element_text(size = 18, color = "black", face = "bold") else element_blank(),  # Larger x-axis text
      axis.text.y.left = if (show_labels) element_text(size = 18, color = "black", face = "bold") else element_blank(),  # Larger left y-axis text
      axis.text.y.right = if (show_labels) element_text(size = 18, color = "black", face = "bold") else element_blank(),  # Larger right y-axis text
      axis.ticks.length = unit(0.25, "cm"),  # Increase tick length slightly
      axis.ticks = element_line(size = 1.2),  # Make ticks visible
      axis.title.x = element_text(size = 20, face = "bold"),  # Larger bold x-axis label
      axis.title.y.left = if (show_labels) element_text(size = 20, face = "bold"),  # Larger bold left y-axis label for VIP Value
      axis.title.y.right = if (show_labels) element_text(size = 20, face = "bold"),  # Larger bold right y-axis label for Spearman's rho
      plot.margin = margin(t = 10, r = 10, b = 10, l = 10)
    )
  
  return(p)
}


# Create and print plots for all unique neighbor IDs in the dataset
for (id in unique(plot_data$ID)) {
  if (id == "PLGY01") {
    plot <- create_vip_correlation_plot(plot_data, id, show_labels = TRUE)
  } else {
    plot <- create_vip_correlation_plot(plot_data, id, show_labels = TRUE)
  }
  print(plot)
}


```

```{r}
library(svglite)
library(Cairo)
# Directory to save the plots
output_dir <- "./Output"

if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}

# Create and save plots as PDF and SVG
for (id in unique(plot_data$ID)) {
  plot <- if (id == "PLGY01") {
    create_vip_correlation_plot(plot_data, id, show_labels = TRUE)
  } else {
    create_vip_correlation_plot(plot_data, id, show_labels = FALSE)
  }
  
  # Define file paths
  pdf_path <- file.path(output_dir, paste0("plot_", id, ".pdf"))
  svg_path <- file.path(output_dir, paste0("plot_", id, ".svg"))
  
  # Save as PDF using cairo_pdf
  ggsave(pdf_path, plot = plot, width = 8, height = 6, device = cairo_pdf, dpi = 300)
  
  # Save as SVG
  ggsave(svg_path, plot = plot, width = 8, height = 6, device = "svg")
}


```




```{r}


# # Define the nodes that should have solid lines
#solid_nodes <- c("PLGY01", "20170700_S25_9", "20170700_S25_26", "20150700_S15_15", 
#                  "20140700_E22_12", "20110800_S1D_5", "20120700_S1X_2")
solid_nodes <- c("3300037175_6","20170700_S25_9","20160700_S15_2","20170700_S25_26","20140700_E22_12","20110800_E3D_9","20110800_S1D_14","20120700_S1X_2", "PLGY01")

# 
# # Create a new column for edge attributes
# edge_attributes <- unique_combined_df %>%
#   mutate(edge_line = ifelse(node_id %in% solid_nodes & neighbor_id %in% solid_nodes, "solid", "dashed"))
# 
# # Rename columns to match Cytoscape's expected format
# edge_attributes <- edge_attributes %>%
#   rename(Source = node_id, Target = neighbor_id)
# 
# # Write the edge attributes to a CSV file
# write.csv(edge_attributes, "edge_attributes.csv", row.names = FALSE, col.names = TRUE, quote = FALSE)

# Filter to keep only connections directly to or from the solid nodes
filtered_df <- unique_combined_df %>%
  filter(node_id %in% solid_nodes | neighbor_id %in% solid_nodes)

# Create a new column for edge attributes
filtered_df <- filtered_df %>%
  mutate(edge_line = ifelse(node_id %in% solid_nodes & neighbor_id %in% solid_nodes, "solid", "dashed"))

# Rename columns to match Cytoscape's expected format
edge_attributes_df <- filtered_df %>%
  rename(Source = node_id, Target = neighbor_id)

# Write the edge attributes to a CSV file
write.csv(edge_attributes_df %>% select(Source, Target, edge_line, num_layers), "edge_attributes.csv", row.names = FALSE, col.names = TRUE, quote = FALSE)

# Print the DataFrame to check the results
print(edge_attributes_df)


```


## Summary
This document provides a detailed analysis of network interactions and their relationships with VIP values. The workflow includes defining functions for network analysis, processing nodes and their neighbors, loading VIP values, and filtering data to highlight important findings.