---
title: "02_wgcna"
author: "Dylan"
date: '2022-05-20'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!dir.exists("output")) {
  dir.create("output", recursive = TRUE)
}
```

#Load data and packages
```{r load-data}

library(ggplot2)
library(phyloseq)
library(ggpubr)
library(dplyr)
library(microbiome)
library(vegan)
library(reshape2)
library(tidyr)
library(stringr)
library(tidyverse)     # tidyverse will pull in ggplot2, readr, other useful libraries
library(WGCNA)    
library(gplots)
library(RColorBrewer)

set.seed(1)

# Set memory_saving to FALSE so setup.R does not delete input_counts_all
memory_saving <- FALSE

# Source setup.R to load all clean datasets and libraries (using cache if available)
library(here)
cache_path <- here("data", "setup_cache.RData")
if (file.exists(cache_path)) {
  load(cache_path)
} else {
  source(here("setup.R"))
  save(input_counts_all, file = cache_path)
}

# 1. Use the input_counts_all$otu_table generated from setup.R
otumat <- as.data.frame(input_counts_all$otu_table)
rownames(otumat) <- otumat$genome
otumat$genome <- NULL
OTU <- otu_table(otumat, taxa_are_rows = TRUE)

# 2. Use the metadata generated from setup.R
md <- as.data.frame(input_counts_all$sample_metadata)
md$metaG_JGI_NovaSeq__ <- md$temporal_sample_id
rownames(md) <- md$metaG_JGI_NovaSeq__

# Clean up / subset metadata: remove columns that aren't traits
cols_to_remove <- c(
  "SampleID__", "SampleID_old__", "Month__", "Core__", "image_url__", "DepthMin__", "metaG_JGI_HiSeq__", "FTIR__", 
  "Weather_Conditions_on_Day_of_Coring", "UpdateDate__.DepthInfo", "Notes__.DepthInfo", "sample", "DepthLumping", 
  "CoreGroup__", "Site__", "GPS__", "DepthCode__", "DepthMax__", "metaG_ACE__", "EEMS__",
  "WTD.cm_neg_is_below_sfc__", "UpdateDate__.Biogeochemistry", "Notes__.Biogeochemistry", "temporal_sample_id", "Date__", "FieldSampling__", "pool.DOM", "pool.dD_CH4", "Phosphate.uM__",
  "sd_AirTemperature_7d", "sd_AirTemperature_21d", "sd_AirTemperature_14d", "sd_AirTemperature_28d",
  "sd_AirTemperature_growing", "sd_AirTemperature_all_growing", "n_AirTemperature_7d",
  "n_AirTemperature_14d", "n_AirTemperature_21d", "n_AirTemperature_28d", "n_AirTemperature_growing",
  "n_AirTemperature_all_growing", "ALD_detectable", "Outlier_status", "DepthID"
)
md <- md[, !(names(md) %in% cols_to_remove), drop = FALSE]

md$Habitat__ <- factor(md$Habitat__)
md$Habitat__ <- factor(md$Habitat__, levels(md$Habitat__)[c(3,1,2)])
md$Year__ <- factor(md$Year__)
sampledata <- sample_data(md)

# 3. Use the taxonomy generated from setup.R
tax <- as.data.frame(input_counts_all$taxonomy)
rownames(tax) <- tax$genome
tax$genome <- NULL
tax <- as.matrix(tax)
TAX <- tax_table(tax)

physeq <- phyloseq(OTU, TAX)
physeq <- merge_phyloseq(physeq, sampledata)
summarize_phyloseq(physeq)

physeq
physeq = filter_taxa(physeq, function(x) sum(x) > 0, TRUE)
physeq



```


```{r wgcna-functions}

prepare_wgcna_data <- function(pseq, min_num_samples) {
    otumat <- as(otu_table(pseq), "matrix")
    otumat <- otumat[,order(colnames(otumat))]
  
    otumat[is.na(otumat)] <- 0
    
    sub <- as.data.frame(apply(otumat, 1, function(x)sum(x > 0)))
    colnames(sub) <- c("Vals")
    otumat <- as.data.frame(otumat)
    otumat$Pops <- row.names(otumat)
    sub <- subset(sub, Vals > min_num_samples)
    otumat <- subset(otumat, Pops %in% row.names(sub))
    otumat$Pops <- NULL
    
    data <- otumat
    data <- t(data)
    cov = decostand((data),method="hellinger")
    gsg = goodSamplesGenes(cov, verbose = 3);
    
    if (!gsg$allOK)
    {
      if (sum(!gsg$goodGenes)>0)
        printFlush(paste("Removing genes:", paste(names(data)[!gsg$goodGenes], collapse = ", ")));
      if (sum(!gsg$goodSamples)>0)
        printFlush(paste("Removing samples:", paste(rownames(data)[!gsg$goodSamples], collapse = ", ")));
      cov = cov[gsg$goodSamples, gsg$goodGenes]
    }
    
    Metadata<-md
    rownames(Metadata)=Metadata$metaG_JGI_NovaSeq__
    cov <- as.data.frame(cov)
    cov$pops <- row.names(cov)
    metadata_sub <- subset(Metadata, metaG_JGI_NovaSeq__ %in% row.names(cov))
    cov_sub <- subset(cov, pops %in% Metadata$metaG_JGI_NovaSeq__)
    Metadata <- metadata_sub
    cov <- cov_sub
    Metadata <- Metadata[order(Metadata$metaG_JGI_NovaSeq__),]
    cov <- cov[order(cov$pops),]
    Metadata$metaG_JGI_NovaSeq__ <- NULL
    cov$pops <- NULL
    
    return(list(cov = cov, Metadata = Metadata))
}

build_wgcna_network <- function(cov, power_cov, habitat, save_outputs = FALSE, plot_dendrogram = FALSE) {
    allowWGCNAThreads()
    
    net = blockwiseModules(cov, power = power_cov,
                           TOMType = "signed", networkType = "signed",
                           reassignThreshold = 0, mergeCutHeight = 0.01,
                           numericLabels = TRUE, pamRespectsDendro = FALSE,
                           saveTOMs = TRUE,
                           saveTOMFileBase = file.path("output", "MAGs_TOM"), 
                           verbose = 3, maxBlockSize = 1800)
    table(net$colors)  
    mergedColors = labels2colors(net$colors)
    
    clusterid <- as.data.frame(net$colors)
    clusteridcolors <- as.data.frame(mergedColors)
    rownames(clusteridcolors) <- rownames(clusterid)
    clustermembership <- merge(clusterid,clusteridcolors, by='row.names')
    row.names(clustermembership) <- clustermembership$Row.names
    clustermembership$Row.names <- NULL
    
    if (save_outputs) {
        write.csv(file=file.path("output", paste0(habitat,"_Module_membership.csv")),clustermembership) 
    }
    
    if (plot_dendrogram) {
        sizeGrWindow(12, 9)
        plotDendroAndColors(net$dendrograms[[1]], mergedColors[net$blockGenes[[1]]],
                            "Module colors",
                            dendroLabels = FALSE, hang = 0.03,
                            addGuide = TRUE, guideHang = 0.05)
    }
    
    moduleLabels = net$colors
    moduleColors = labels2colors(net$colors)
    geneTree = net$dendrograms[[1]];
    
    MAGs_eig = moduleEigengenes(cov, moduleColors)$eigengenes
    pops = orderMEs(MAGs_eig)
    rownames(pops) <- rownames(cov)
    
    return(list(net = net, pops = pops, moduleColors = moduleColors, moduleLabels = moduleLabels, geneTree = geneTree))
}

run_wgcna <- function(pseq, min_num_samples, power_cov, habitat) {
    prep <- prepare_wgcna_data(pseq, min_num_samples)
    net_res <- build_wgcna_network(prep$cov, power_cov, habitat, save_outputs = TRUE, plot_dendrogram = TRUE)
    pops <- net_res$pops
    moduleLabels <- net_res$moduleLabels
    moduleColors <- net_res$moduleColors
    geneTree <- net_res$geneTree
    save(pops, moduleLabels, moduleColors, geneTree, file = file.path("output", "MAGs_Module-networkConstruction-auto.RData"))
    write.csv(file=file.path("output", paste0(habitat,".04.MAGs_Module_eigen_values.csv")), pops)
    
    nSamples = nrow(prep$cov)
    moduleTraitCor = cor(net_res$pops, prep$Metadata, use = "p")
    moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples)
    
    combined_table <- as.data.frame(matrix(nrow = nrow(moduleTraitCor), ncol = ncol(moduleTraitCor)))
    colnames(combined_table) <- colnames(moduleTraitCor)
    rownames(combined_table) <- rownames(moduleTraitCor)
    
    for (col in colnames(moduleTraitPvalue)) {
      combined_table[[col]] <- paste0(
        signif(moduleTraitCor[, col], 2), 
        " (p=", 
        signif(moduleTraitPvalue[, col], 2), 
        ")"
      )
    }
    write.csv(combined_table, file = file.path("output", paste0(habitat, "_Combined_ModuleTrait_Table.csv")), row.names = TRUE)

    # Generate long-form table with multiple testing correction (BH)
    p_vector <- as.vector(moduleTraitPvalue)
    q_values <- p.adjust(p_vector, method = "BH", n = sum(!is.na(p_vector)))
    modules <- rownames(moduleTraitCor)
    traits <- colnames(moduleTraitCor)
    
    long_form <- data.frame(
        Module = rep(modules, times = length(traits)),
        Habitat = habitat,
        Metadata_Parameter = rep(traits, each = length(modules)),
        Correlation = as.vector(moduleTraitCor),
        P_value = as.vector(moduleTraitPvalue),
        Q_value = q_values,
        stringsAsFactors = FALSE
    )
    write.csv(long_form, file = file.path("output", paste0(habitat, "_LongForm_ModuleTrait_Table.csv")), row.names = FALSE)

    textMatrix = paste(signif(moduleTraitCor, 2), "\n(",signif(moduleTraitPvalue, 1), ")", sep = "")
    dim(textMatrix) = dim(moduleTraitCor)
    library(RColorBrewer)
    pdf(file.path("output", paste0(habitat,".05.Module-trait relationships.pdf")),width=20,height=14)
    par(mar=c(8,12,4,5))
    labeledHeatmap(Matrix = moduleTraitCor,
                   xLabels = names(prep$Metadata),
                   yLabels = names(net_res$pops),
                   ySymbols = substring(names(net_res$pops), 3),
                   colorLabels = FALSE,
                   colors = brewer.pal(11,"RdBu"),
                   textMatrix = textMatrix,
                   setStdMargins = FALSE,
                   cex.text = .5,
                   cex.lab = 1.2,
                   zlim = c(-1,1),
                   main = paste("Module-trait relationships"))
    dev.off()
}

run_spls <- function(pseq, min_num_samples, power_cov, habitat, module, parameter) {
    prep <- prepare_wgcna_data(pseq, min_num_samples)
    net_res <- build_wgcna_network(prep$cov, power_cov, habitat, save_outputs = FALSE, plot_dendrogram = FALSE)
    
    nSamples = nrow(prep$cov)
    moduleTraitCor = cor(net_res$pops, prep$Metadata, use = "p")
    moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples)
    textMatrix = paste(signif(moduleTraitCor, 2), "\n(",signif(moduleTraitPvalue, 1), ")", sep = "")
    dim(textMatrix) = dim(moduleTraitCor)

    weight <- as.data.frame(prep$Metadata[,parameter])
    weight_t<- as.data.frame(weight)
    rownames(weight_t) <- rownames(prep$cov)
    
    weight_t <-na.omit(weight_t)
    cleannamesweight <- rownames(weight_t)
    names(weight_t) = parameter
    weight <- as.data.frame(weight_t)
    MAGs_clean <- net_res$pops[rownames(weight_t),]
    cov_clean <- prep$cov[rownames(weight_t),]
    
    modNames = substring(names(net_res$pops), 3)
    geneModuleMembership = as.data.frame(cor(prep$cov, net_res$pops, use = "p"));
    MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));
    names(geneModuleMembership) = paste("MM", modNames, sep="");
    names(MMPvalue) = paste("p.MM", modNames, sep="");
    geneTraitSignificance = as.data.frame(cor(cov_clean, weight, use = "p"));
    GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));
    names(geneTraitSignificance) = paste("GS.", names(weight), sep="");
    names(GSPvalue) = paste("p.GS.", names(weight), sep="");
    
    column = match(module, modNames);
    moduleGenes = net_res$moduleColors==module
    pdf(file.path("output", paste0(habitat,".06.", module,"Module membership vs. population significance for", parameter,".pdf")),width=12,height=8)
    par(mfrow=c(1,1))
    par(mar=c(8,8,8,8))
    verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]),
                       abs(geneTraitSignificance[moduleGenes, 1]),
                       xlab = paste("Module Membership -", module),
                       ylab = paste("Correlation to",parameter), lmFnc = lm,
                       abline = TRUE, abline.color = 1, abline.lty = 1, 
                       corFnc = "cor", corOptions = "use = 'p'", 
                       main = paste("\n"),
                       cex.main = 2, cex.lab = 2, cex.axis = 1.2, pch=21, bg=module, col="black")
    dev.off()

    MET = orderMEs(cbind(MAGs_clean, weight))
    pdf(file.path("output", paste0(habitat,".07.Adjacency_trait_ME for", parameter,".pdf")))
    par(cex = 1.0)
    plotEigengeneNetworks(MET, paste("Eigengene dendrogram with ",parameter), marDendro = c(0,4,2,0),plotHeatmaps = FALSE)
    par(cex = 1.0)
    plotEigengeneNetworks(MET, paste("Eigengene adjacency heatmap with ",parameter), marHeatmap = c(3,4,2,2),plotDendrograms = FALSE, xLabelsAngle = 90)
    dev.off()
    
    th_r2<-0.3 
    subnetwork<-cov_clean[,moduleGenes]
    subnetwork<- as.matrix(subnetwork)
    
    library("pls")
    pls_result<-plsr(as.matrix(weight) ~ (subnetwork), validation="LOO",method="oscorespls")
    r2_vector<-R2(pls_result)
    max<-0
    max_comp<--1
    for (j in 1:length(r2_vector$val)){
      if(r2_vector$val[j]>th_r2){
        if(r2_vector$val[j]>max){
          max<-r2_vector$val[j]
          max_comp<-r2_vector$comp[j]
        }
      }
    }
    
    source("http://mevik.net/work/software/VIP.R")
    library(ggplot2)
    print(paste(" the max r2 is ",max," corresponding to comp ",max_comp,sep=""))
    if(max==0){
      print ("No good correlation, we stop here")
      df_3 <- NULL
    }else {
      print("Good correlation, we check the VIP !")
      output<-file.path("output", paste0(habitat,".08.VIP_values_with_",parameter,module,".txt"))
      vip_result<-VIP(pls_result)
      vip_components<-sort(vip_result[max_comp,],decreasing=TRUE)[1:100]
      for (i in 1:100){
        cat(paste("Rank ",i," we have ",names(vip_components[i])," with a VIP of ",vip_components[i],"\n",sep=""),file=output,append=TRUE)
      }
      weight_2 <- as.data.frame(weight[!is.na(weight)])
      rownames(weight_2) <-cleannamesweight
      df<-data.frame(x=weight_2,y=pls_result$validation$pred[,,max_comp])
      colnames(df)<-c("x","y")
      cor.test(df$x,df$y)
      df_2<-as.data.frame(as.double(cor(subnetwork, weight, use = "p")))
      colnames(df_2)<-c("Radius")
      rownames(df_2)<-colnames(subnetwork)
      write.csv(df_2,file=file.path("output", paste0(habitat,".09.Normal_correlation",module, parameter,".txt")))
      df_2<-cbind(Name=row.names(df_2),df_2)
      table<-row.names(df_2)
      df_2<-cbind(df_2,Axis=as.integer(c(1)))
      df_3<-as.data.frame(as.double(vip_result[max_comp,]))
      rownames(df_3)<-colnames(subnetwork)
      colnames(df_3)<-c("Radius")
      write.csv(df_3,file=file.path("output", paste0(habitat,".10.VIP_correlation",module, parameter,".txt")))
      
      lmp <- function (modelobject) {
        if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
        f <- summary(modelobject)$fstatistic
        p <- pf(f[1],f[2],f[3],lower.tail=F)
        attributes(p) <- NULL
        return(p)
      }
      
      regression <- lm(df$x~df$y)
      pdf(file.path("output", paste0(habitat,".11.measured_vs_predicted_",module,"-vs-",parameter,".pdf")))
      print(ggplot(data=df) + theme_bw() + geom_abline() + geom_smooth(aes(x=x,y=y),method=lm, se = FALSE) + 
        xlab("Measured") + ylab("Predicted") + ggtitle(paste("Comparison of",parameter,"measured vs predicted for module",module)) + 
        theme(plot.title = element_text(hjust = 0.5),axis.text=element_text(color="black",size=10),axis.title=element_text(color="black",size=15), axis.ticks=element_line(color="black"),panel.grid.major = element_blank(),panel.grid.minor = element_blank())+ 
        geom_point(aes(x=x,y=y), size=2.5, shape=21, fill=module, color="black") +
        scale_x_continuous(breaks = pretty(df$x, n=5)) +
        scale_y_continuous(breaks = pretty(df$x, n=5)) +
        annotate("text",x=Inf,y=-Inf,hjust=1.1,vjust=-0.5,label= paste("r2 =",format(round(summary(regression)$r.squared, 2), nsmall = 2) , "\n", "p-value =", format(round(lmp(regression), 2), nsmall = 2))))
      dev.off()
    }
    
    return(list(geneModuleMembership = geneModuleMembership, geneTraitSignificance = geneTraitSignificance, VIPtable = df_3))
}

determine_thresholding <- function(pseq, min_num_samples) {
    prep <- prepare_wgcna_data(pseq, min_num_samples)
    
    sampleTree = hclust(dist(prep$cov), method = "average");
    sizeGrWindow(12,9)
    par(cex = 0.6);
    par(mar = c(10,4,4,2))
    plot(sampleTree, main = "Sample clustering to detect outliers", sub="", xlab="", cex.lab = 1.5,
    cex.axis = 1.5, cex.main = 2)

    input_mat = prep$cov
    allowWGCNAThreads()
    powers = c(c(1:10), seq(from = 12, to = 20, by = 2))
    sft = pickSoftThreshold(
      input_mat,
      powerVector = powers,
      verbose = 5,
      networkType = "signed"
      )
    
    par(mfrow = c(1,2));
    par(mar = c(5,5,4,2));
    cex1 = 0.9;
    
    plot(sft$fitIndices[, 1],
         -sign(sft$fitIndices[, 3]) * sft$fitIndices[, 2],
         xlab = "Soft Threshold (power)",
         ylab = "Scale Free Topology Model Fit, signed R^2",
         main = paste0("Scale independence - ", min_num_samples)
    )
    text(sft$fitIndices[, 1],
         -sign(sft$fitIndices[, 3]) * sft$fitIndices[, 2],
         labels = powers, cex = cex1, col = "red"
    )
    abline(h = 0.80, col = "red")
    plot(sft$fitIndices[, 1],
         sft$fitIndices[, 5],
         xlab = "Soft Threshold (power)",
         ylab = "Mean Connectivity",
         type = "n",
         main = paste0("Mean connectivity - ", min_num_samples)
    )
    text(sft$fitIndices[, 1],
         sft$fitIndices[, 5],
         labels = powers,
         cex = cex1, col = "red")
  
}

```{r palsa-thresholding}
pseq <- subset_samples(physeq, (Habitat__ == "Palsa"))
pdf(file.path("output", "Palsa_Thresholding.pdf"), width = 12, height = 9)
for (i in 1:20) {
  print(i)
  determine_thresholding(pseq, i)
}
dev.off()
```

```{r palsa-wgcna}

pseq <- subset_samples(physeq, (Habitat__ == "Palsa"))
run_wgcna(pseq, min_num_samples = 10, power_cov = 14, "Palsa")

```

```{r bog-thresholding}
pseq <- subset_samples(physeq, (Habitat__ == "Bog"))
pdf(file.path("output", "Bog_Thresholding.pdf"), width = 12, height = 9)
for (i in 1:20) {
  print(i)
  determine_thresholding(pseq, i)
}
dev.off()
```

```{r bog-wgcna}

pseq <- subset_samples(physeq, (Habitat__ == "Bog"))
run_wgcna(pseq, min_num_samples = 1, power_cov = 14, "Bog")
vip_df = run_spls(pseq,  min_num_samples = 1, power_cov = 14, "Bog", module = "blue", parameter = "alphaC")
vip_df = run_spls(pseq,  min_num_samples = 1, power_cov = 14, "Bog", module = "blue", parameter = "T_soil.deg_C")
```


```{r fen-thresholding}
pseq <- subset_samples(physeq, (Habitat__ == "Fen"))
pdf(file.path("output", "Fen_Thresholding.pdf"), width = 12, height = 9)
for (i in 1:20) {
   print(i)
   determine_thresholding(pseq, i)
}
dev.off()
```

```{r fen-wgcna}

pseq <- subset_samples(physeq, (Habitat__ == "Fen"))
run_wgcna(pseq, min_num_samples = 6, power_cov = 12, "Fen")

```

```{r all-thresholding}
pdf(file.path("output", "AllSamples_Thresholding.pdf"), width = 12, height = 9)
for (i in 1:20) {
  print(i)
  determine_thresholding(physeq, i)
}
dev.off()
```


```{r all-wgcna}

run_wgcna(physeq, min_num_samples = 9, power_cov = 16, "All")
vip_df = run_spls(physeq,  min_num_samples = 9, power_cov = 16, "All", module = "lightgreen", parameter = "alphaC")

# vip_df = run_spls(physeq,  min_num_samples = 9, power_cov = 16, "All", module = "lightgreen", parameter = "d13C_CO2__")
# #vip_df = run_spls(physeq,  min_num_samples = 9, power_cov = 16, "All", module = "lightgreen", parameter = "CO2.mM__")
# ##vip_df = run_spls(physeq,  min_num_samples = 9, power_cov = 16, "All", module = "brown", parameter = "DOC.mM__")
# #vip_df = run_spls(physeq,  min_num_samples = 9, power_cov = 16, "All", module = "brown", parameter = "TN.mM__")

```

# <!-- ```
# ```{r}

# physeq_genus = tax_glom(physeq, "Genus")

# for (i in 1:20) {
#   print(i)
#   determine_thresholding(physeq_genus, i)
# }

# run_wgcna(physeq_genus,1,7,"Genus")

# run_spls(physeq_genus,1,7,"Genus","red","alphaC")
# run_spls(physeq_genus,1,7,"Genus","red","DOC.mM__")
# run_spls(physeq_genus,1,7,"Genus","red","TN.mM__")
# run_spls(physeq_genus,1,7,"Genus","pink","Sulfate.uM__")
# run_spls(physeq_genus,1,7,"Genus","turquoise","d13C_CH4__")
# ```

# ```{r}

# physeq_genus_palsa = subset_samples(physeq_genus, Habitat__ == "Palsa")

# for (i in 1:20) {
#   print(i)
#   determine_thresholding(physeq_genus_palsa, i)
# }

# run_wgcna(physeq_genus_palsa,1,7,"Genus.Palsa")
# run_spls(physeq_genus_palsa,1,7,"Genus.Palsa","turquoise","d13C_CO2__")
# run_spls(physeq_genus_palsa,1,7,"Genus.Palsa","turquoise","T_soil.deg_C")
# run_spls(physeq_genus_palsa,1,7,"Genus.Palsa","green","CO2.percent") #no good correlation

# ```

# ```{r}

# physeq_genus_bog = subset_samples(physeq_genus, Habitat__ == "Bog")

# for (i in 1:20) {
#   print(i)
#   determine_thresholding(physeq_genus_bog, i)
# }

# run_wgcna(physeq_genus_bog,3,8,"Genus.Bog")

# run_spls(physeq_genus_bog,3,8,"Genus.Bog","blue","N.percent__")
# run_spls(physeq_genus_bog,3,8,"Genus.Bog","blue","d15N_peat__")
# run_spls(physeq_genus_bog,3,8,"Genus.Bog","blue","DepthAvg__")

# ```
# ```{r}

# physeq_genus_fen = subset_samples(physeq_genus, Habitat__ == "Fen")

# for (i in 1:20) {
#   print(i)
#   determine_thresholding(physeq_genus_fen, i)
# }

# run_wgcna(physeq_genus_fen,1,12,"Genus.Fen")

# run_spls(physeq_genus_fen,3,8,"Genus.Bog","blue","N.percent__")
# #run_spls(physeq_genus_bog,3,8,"Genus.Bog","blue","d15N_peat__")
# #run_spls(physeq_genus_bog,3,8,"Genus.Bog","blue","DepthAvg__")

# ```


# ```{r}
# for (i in 1:20) {
#   print(i)
#   determine_thresholding(physeq, i)
# }


# ```


# ```{r}

# run_wgcna(physeq, min_num_samples = 9, power_cov = 16, "All")


# ```

# ```{r}


# #browser()
# #vip_df = run_spls(physeq,  min_num_samples = 9, power_cov = 16, "All", module = "lightgreen", parameter = "alphaC")
# vip_df = run_spls(physeq,  min_num_samples = 9, power_cov = 16, "All", module = "lightgreen", parameter = "d13C_CO2__")
# #vip_df = run_spls(physeq,  min_num_samples = 9, power_cov = 16, "All", module = "lightgreen", parameter = "CO2.mM__")
# ##vip_df = run_spls(physeq,  min_num_samples = 9, power_cov = 16, "All", module = "brown", parameter = "DOC.mM__")
# #vip_df = run_spls(physeq,  min_num_samples = 9, power_cov = 16, "All", module = "brown", parameter = "TN.mM__")
# #vip_df = run_spls(physeq,  min_num_samples = 9, power_cov = 16, "All", module = "lightcyan", parameter = "pH_peat") #did not work
# #vip_df = run_spls(physeq,  min_num_samples = 9, power_cov = 16, "All", module = "salmon", parameter = "CO2.percent") #did not work
# #vip_df = run_spls(physeq,  min_num_samples = 9, power_cov = 16, "All", module = "turquoise", parameter = "d13C_CH4__")



# vip_df <- subset(vip_df, Radius > 1)
# temp2 <- merge(vip_df, tax, by = "row.names")

# ```

# ```{r}
# pseq <- subset_samples(physeq, (Habitat__ == "Palsa"))
# for (i in 1:20) {
#   print(i)
#   determine_thresholding(pseq, i)
# }

# ```

# ```{r}

# pseq <- subset_samples(physeq, (Habitat__ == "Palsa"))
# run_wgcna(pseq, min_num_samples = 10, power_cov = 14, "Palsa")

# ```


# ```{r}


# #vip_df = run_spls(pseq,  min_num_samples = 10, power_cov = 14, "Palsa", module = "blue", parameter = "d15N_peat__") #significant but ugly
# #vip_df = run_spls(pseq,  min_num_samples = 10, power_cov = 14, "Palsa", module = "yellow", parameter = "C.percent__") #significant but ugly


# #vip_df <- subset(vip_df, Radius > 1)
# #temp2 <- merge(vip_df, tax, by = "row.names")

# colors <- c("yellow","brown","red", "turquoise")#, "pink") #turquoise, red, pink

# temp_lineages <- list()

# for (c in colors){
#   vip_df = run_spls(pseq,  min_num_samples = 10, power_cov = 14, "Palsa", module = c, parameter = "T_soil.deg_C")
#   temp <- subset(vip_df, Radius > 1)
#   temp_lineages[[c]] <- row.names(temp)
# }

# ```
# ```{r}

# test <- plyr::ldply(temp_lineages, rbind)
# row.names(test) <- test$.id
# test$.id <- NULL
# test<- t(test)
# test_palsa <- melt(test)
# test_palsa$Habitat <- "Palsa"
# test_palsa <- na.omit(test_palsa)
# test_palsa$Var1 <- NULL
# colnames(test_palsa) <- c("module","mag","habitat")
# #write.csv(test,file="palsa_lineages_cor_temp.txt",quote=FALSE, row.names = FALSE)

# ```

# ```{r}
# pseq <- subset_samples(physeq, (Habitat__ == "Bog"))
# #run_wgcna(pseq, 8, 9, "Palsa")
# for (i in 1:20) {
#   print(i)
#   determine_thresholding(pseq, i)
# }

# ```



# ```{r}

# pseq <- subset_samples(physeq, (Habitat__ == "Bog"))
# run_wgcna(pseq, min_num_samples = 1, power_cov = 14, "Bog")

# ```

# ```{r}



# #vip_df = run_spls(pseq,  min_num_samples = 1, power_cov = 14, "Bog", module = "blue", parameter = "N.percent__") #significant but not much data
# #vip_df = run_spls(pseq,  min_num_samples = 1, power_cov = 14, "Bog", module = "blue", parameter = "CO2.mM__")
# #vip_df = run_spls(pseq,  min_num_samples = 1, power_cov = 14, "Bog", module = "green", parameter = "dD_CH4__") #no good correlation -- from function output message
# vip_df = run_spls(pseq,  min_num_samples = 1, power_cov = 14, "Bog", module = "blue", parameter = "alphaC")
# #vip_df = run_spls(pseq,  min_num_samples = 1, power_cov = 14, "Bog", module = "blue", parameter = "d13C_CO2__")


# #vip_df <- subset(vip_df, Radius > 1)
# #temp2 <- merge(vip_df, tax, by = "row.names")

# colors <- c("blue")#, "pink") #turquoise, red, pink

# temp_lineages_bog <- list()

# for (c in colors){
#   vip_df = run_spls(pseq,  min_num_samples = 1, power_cov = 14, "Bog", module = c, parameter = "T_soil.deg_C")
#   temp <- subset(vip_df, Radius > 1)
#   temp_lineages_bog[[c]] <- row.names(temp)
# }

# ```

# ```{r}

# test <- plyr::ldply(temp_lineages_bog, rbind)
# row.names(test) <- test$.id
# test$.id <- NULL
# test<- t(test)
# test_bog <- melt(test)
# test_bog$Habitat <- "Bog"
# test_bog <- na.omit(test_bog)
# test_bog$Var1 <- NULL
# colnames(test_bog) <- c("module","mag","habitat")
# #write.csv(test,file="bog_lineages_cor_temp.txt",quote=FALSE, row.names = FALSE)

# ```


# ```{r}
# pseq <- subset_samples(physeq, (Habitat__ == "Fen"))
# #run_wgcna(pseq, 8, 9, "Palsa")
# for (i in 1:20) {
#   print(i)
#   determine_thresholding(pseq, i)
# }

# ```

# ```{r}

# pseq <- subset_samples(physeq, (Habitat__ == "Fen"))
# run_wgcna(pseq, min_num_samples = 6, power_cov = 12, "Fen")

# ```

# ```{r}


# #vip_df = run_spls(pseq,  min_num_samples = 6, power_cov = 12, "Fen", module = "brown", parameter = "median_AirTemperature_14d")
# #vip_df = run_spls(pseq,  min_num_samples = 6, power_cov = 12, "Fen", module = "pink", parameter = "d13C_peat__")
# #vip_df = run_spls(pseq,  min_num_samples = 6, power_cov = 12, "Fen", module = "blue", parameter = "N.percent__")
# #vip_df = run_spls(pseq,  min_num_samples = 6, power_cov = 12, "Fen", module = "blue", parameter = "d15N_peat__")
# vip_df = run_spls(pseq,  min_num_samples = 6, power_cov = 12, "Fen", module = "yellow", parameter = "C.percent__")
# pseq = subset_samples(physeq, (Habitat__ == "Fen" & median_AirTemperature_14d < 14))
# vip_df = run_spls(pseq,  min_num_samples = 6, power_cov = 12, "Fen", module = "brown", parameter = "median_AirTemperature_14d")

# #vip_df <- subset(vip_df, Radius > 1)
# #temp2 <- merge(vip_df, tax, by = "row.names")

# colors <- c("turquoise")#, "pink") #turquoise, red, pink

# temp_lineages_fen <- list()

# for (c in colors){
#   vip_df = run_spls(pseq,  min_num_samples = 6, power_cov = 12, "Fen", module = c, parameter = "T_soil.deg_C")
#   temp <- subset(vip_df, Radius > 1)
#   temp_lineages_fen[[c]] <- row.names(temp)
# }

# colors <- c("brown")#, "pink") #turquoise, red, pink


# for (c in colors){
#   vip_df = run_spls(pseq,  min_num_samples = 6, power_cov = 12, "Fen", module = c, parameter = "median_AirTemperature_14d")
#   temp <- subset(vip_df, Radius > 1.5)
#   temp_lineages_fen[[c]] <- row.names(temp)
# }

# ```

# ```{r}
# test <- plyr::ldply(temp_lineages_fen, rbind)
# row.names(test) <- test$.id
# test$.id <- NULL
# test<- t(test)
# test_fen <- melt(test)
# test_fen$Habitat <- "Fen"
# test_fen <- na.omit(test_fen)
# test_fen$Var1 <- NULL
# colnames(test_fen) <- c("module","mag","habitat")
# #write.csv(test,file="fen_lineages_cor_temp.txt",quote=FALSE, row.names = FALSE)

# ```


# #Shallow samples
# ```{r}
# pseq <- subset_samples(physeq, (Habitat__ == "Bog" & (DepthAvg__ < 20)))
# #run_wgcna(pseq, 8, 9, "Palsa")
# for (i in 1:20) {
#   print(i)
#   determine_thresholding(pseq, i)
# }

# ```

# ```{r}

# run_wgcna(pseq, min_num_samples = 2, power_cov = 12, "Bog-shallow")

# ```

# ```{r}


# vip_df = run_spls(pseq, min_num_samples = 18, power_cov = 8, "Fen", module = "brown", parameter = "median_AirTemperature_14d")
# #vip_df = run_spls(pseq,  min_num_samples = 6, power_cov = 12, "Fen", module = "pink", parameter = "d13C_peat__")
# #vip_df = run_spls(pseq,  min_num_samples = 6, power_cov = 12, "Fen", module = "blue", parameter = "N.percent__")
# #vip_df = run_spls(pseq,  min_num_samples = 6, power_cov = 12, "Fen", module = "blue", parameter = "d15N_peat__")
# #vip_df = run_spls(pseq,  min_num_samples = 6, power_cov = 12, "Fen", module = "yellow", parameter = "C.percent__")


# vip_df <- subset(vip_df, Radius > 1)
# temp2 <- merge(vip_df, tax, by = "row.names")

# ```


# ```{r}

# library(ggpubr)

# ```


# ```{r}

# temp <- subset(md, select = c("Habitat__","alphaC"))
# temp <- na.omit(temp) 
# print(ggplot(temp, aes(x = alphaC, color=Habitat__)) + geom_dotplot())


# temp <- subset(md, select = c("Habitat__","d13C_CO2__"))
# temp <- na.omit(temp) 
# print(ggplot(temp, aes(x = d13C_CO2__, color=Habitat__)) + geom_dotplot())
# ggboxplot(temp, x = "Habitat__", y = "d13C_CO2__") + stat_compare_means() + geom_jitter()


# temp <- subset(md, select = c("Habitat__","CO2.mM__"))
# temp <- na.omit(temp) 
# print(ggplot(temp, aes(x = CO2.mM__, color=Habitat__)) + geom_dotplot())

# temp <- subset(md, select = c("Habitat__","TN.mM__"))
# temp <- na.omit(temp) 
# print(ggplot(temp, aes(x = TN.mM__, color=Habitat__)) + geom_dotplot())

# temp <- subset(md, select = c("Habitat__","DOC.mM__"))
# temp <- na.omit(temp) 
# print(ggplot(temp, aes(x = DOC.mM__, color=Habitat__)) + geom_dotplot())

# temp <- subset(md, select = c("Habitat__","d13C_CH4__"))
# temp <- na.omit(temp) 
# print(ggplot(temp, aes(x = d13C_CH4__, color=Habitat__)) + geom_dotplot())


# ```